The Direction of Search for a Forest Based Bioeconomy

### 2023-09-19 docs/reports/paper/paper1.qmd
add trend and heatmap figure

# set all text on plots to be monospace with rcParams
plt.rcParams["font.family"] = "monospace"

sns.set_theme(style="whitegrid")

```
# Debate around Bioeconomies

@@

##

@@

```{python}
# | label: Queries

@@

```{python}
#| label: Number of Results

display
```

@@

```sql
SELECT *
FROM table
WHERE
    (
        us.use_sectors LIKE '02%' -- Forestry and related services
-- Furniture manufacturing; other manufacturing
        OR us.use_sectors LIKE '20%' -- Wood and wood product manufacturing except furniture
        OR us.use_sectors LIKE '21%' -- Pulp, paper and paper product manufacturing
        OR us.use_sectors LIKE '36%' -- Furniture manufacturing; other manufacturing
        us.use_sectors LIKE '02%' -- Forestry and related services
-- Furniture manufacturing; other manufacturing
        OR product_code LIKE '20%' -- Wood and wood product manufacturing except furniture
        OR product_code LIKE '21%' -- Pulp, paper and paper product manufacturing
        OR product_code LIKE '36%' -- Furniture manufacturing; other manufacturing
    )
    OR (
        description LIKE '%virke%'
        OR description LIKE '%cellulos%'
        OR description LIKE '%lignin%'
        OR description LIKE '%spån%'
        OR description LIKE '%bark%'
        OR description LIKE '%levulinsyra%'
        OR description LIKE '%furfural%'
        OR description LIKE '%svarttjära%'
        OR description LIKE '%svartlut%'
        OR description LIKE '%växtbas%'
        OR description LIKE '%ved%'
        OR description LIKE '%trä%'
        OR description LIKE '%skog%'
        OR description LIKE '%biobränsle%'
        OR description LIKE '%biologisk%'
        OR description LIKE '%nedbrytbar%'
        OR description LIKE '%papper%'
        OR description LIKE '%pappret%'
        OR description LIKE '%karton%'
        OR description LIKE '%tencel%'
    );
```

>>>>

# set all text on plots to be monospace with rcParams
# plt.rcParams["font.family"] = "Ubunto Mono"

sns.set_theme(style="white")

# Debate around Bioeconomies

@@

@@

```{python}
# | label: Query bioeconomy

@@

```{python}
#| label: Query eco-innovation

eco_innovations = pd.read_sql(
    """
SELECT 
    sinno_id
    , innovation_type
FROM
  eco_innovations;
""",
    conn,
)

```{python}
# | label: Query bioeconomy visions

bioeconomy_visions = pd.read_sql(
    """
SELECT bv.sinno_id
, codes.Category AS bioeconomy_vision
, i.year_of_commercialization AS "year"
FROM bioeconomy_visions AS bv
JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
JOIN innovation AS i ON i.sinno_id = bv.sinno_id
WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
UNION
SELECT DISTINCT(ei.sinno_id)
, 'Bioecology Vision' AS bioeconomy_vision
, i.year_of_commercialization AS "i.year"
FROM eco_innovations AS ei
JOIN innovation AS i ON i.sinno_id = ei.sinno_id
WHERE ei.innovation_type IN (206, 107)
AND ei.sinno_id NOT IN (
  SELECT ei2.sinno_id
  FROM eco_innovations as ei2
  WHERE ei2.innovation_type IN (999, '000')
  );
-- adds recycling innovation to bioecology
""",
    conn,
)

bioeconomy_visions.loc[
    bioeconomy_visions["bioeconomy_vision"] == "Not Bioeconomy",
    ["bioeconomy_vision"],
] = "No Bioeconomy Vision"

uncertain_bioeconomy_visions = bioeconomy_visions.query("bioeconomy_vision == 'Unsure'")

certain_bioeconomy_visions = bioeconomy_visions.loc[
    ~bioeconomy_visions["sinno_id"].isin(uncertain_bioeconomy_visions["sinno_id"]), :
]

```{python}
# | label: Number of Results
# | output: asis

display(
    Markdown(
        f"So far, {certain_bioeconomy_visions['sinno_id'].nunique()} innovation have been classified for the period between {certain_bioeconomy_visions['year'].min()} and {certain_bioeconomy_visions['year'].max()}."
    ))
```

```{python}
# | label: fig-bioeconomy_trends
# | fig-cap: Bioeconomy Vision Trends

count_certain_bioeconomy_visions = (
    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
    .size()
    .reset_index(name="count")
)

def plot_bv_trends(df, title):
    palette = sns.color_palette("colorblind")[:5]

    # Create a custom palette with a default light gray color for "not bio" hue
    custom_palette = {
        hue: color if bioeconomy_vision != "No Bioeconomy Vision" else (0.8, 0.8, 0.8)
        for hue, color, bioeconomy_vision in zip(
            df["bioeconomy_vision"], palette, df["bioeconomy_vision"]
        )
    }

    chart = sns.lineplot(
        data=df,
        x="year",
        y="count",
        hue="bioeconomy_vision",
        palette=custom_palette,
        linewidth=2.5,
    )

    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "normal"})
    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "normal"})
    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "normal"})
    chart.tick_params(labelsize=10)

    # Customize x-axis tick labels
    years = df["year"].unique()
    chart.set_xticks(years[::5])
    chart.set_xticklabels(years[::5], rotation=45, ha="right")

    # Create a custom legend with colored labels (no lines)
    handles, labels = chart.get_legend_handles_labels()

    # Set font color for legend labels based on the hue
    legend = chart.legend(
        bbox_to_anchor=(0.5, -0.35), loc="upper center", ncol=2, frameon=False
    )
    for text, label in zip(legend.texts, labels):
        text.set_color(custom_palette[label])

    plt.tight_layout()

plot_bv_trends(count_certain_bioeconomy_visions, title="")

plt.show()
```

```{python}
# | label: fig-vision_heatmap
# | fig-caption: Sectors Contribution to Bioeconomy Visions

sectors = pd.read_sql(
    """
SELECT
  bv.sinno_id as sinno_id,
  c.Category AS vision,
  sni_codes.label AS sector
FROM
  bioeconomy_visions AS bv
  JOIN use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
  JOIN sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
  JOIN classification_codes AS c ON bv.bioeconomy_vision = c.Code;
""",
    conn,
)

sectors = sectors.loc[
    sectors["sinno_id"].isin(set(certain_bioeconomy_visions["sinno_id"])), :
]

sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "No Bioeconomy Vision"

sector_counts = (
    sectors.groupby(["sector", "vision"]).size().to_frame("count").reset_index()
)
# Question is, should I weight this following the way in the env-innovation paper?

sector_counts = sector_counts.sort_values("count", ascending=False)

heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")

plot_heatmap(heatmap_data, "")

# TODO: Better y axis labels

### How many eco-innovations are there?
#### How many by bioeconomy vision?

@@
